---
title: "ReproducibleResearchAssignment1"
author: "siva"
date: "06/14/2015"
output: html_document
---
This markdown file contains the details of the work done to complete assignment 1 of the Reproducible Research course. The assignment requires us to perform some basic analysis on  real-time  measurements from devices such as fitbit, Nike fuelband etc.
This markdown file assumes that the data file is in the same directory.

```{r}
activity_data <- read.csv("activity.csv")
#Saving for posterity
act_data1 <- activity_data
act_data_no_na <- na.omit(activity_data)
str(activity_data)
```

Now I calculate the daily steps
```{r}
daily_steps <-aggregate(act_data_no_na$steps, list(dates=act_data_no_na$date), sum)
```

```{r}
hist(daily_steps$x, breaks=8, col="orange", border="brown")
```

Five-number summary of data
```{r}
#I am not sure if this is less efficient than just calculating mean and median of the steps attribute
summary(daily_steps)
```

Calculate 5 min average
```{r}
five_min_avg <- aggregate(act_data_no_na[,1],  by=list(act_data_no_na$interval), FUN=mean)
colnames(five_min_avg) <- c("Interval", "AvgSteps")
```

```{r}
plot(five_min_avg, type="l", xlab="interval", ylab="average steps", main="Avg. Number of steps per 5-min interval")
```

Which interval has the maximum average number of steps?
```{r}
five_min_avg[five_min_avg[,2]==max(five_min_avg$AvgSteps),][1]
```

Now I calculate the number of rows with NA in the data set
```{r}
sum(is.na(activity_data$steps))
```

Replace NAs with 5 minute average values 
```{r}
for(i in 1:nrow(act_data1)) {
  if(is.na(act_data1[i,"steps"])) act_data1[i,"steps"] <- (five_min_avg[five_min_avg["Interval"]==act_data1[i,"interval"], "AvgSteps"])
  }

str(act_data1)
```



Recalculating with adjusted NA values
```{r}
daily_steps <-aggregate(act_data1$steps, list(dates=act_data1$date), sum)
#plot(daily_steps$dates, daily_steps$x, type="p")
#plot(daily_steps$x,  type="h", main="Histogram of daily steps", xlab="days since beginning", ylab="Number of Steps")
```

```{r}
hist(daily_steps$x, breaks=8, col="lightblue", border="darkblue")
summary(daily_steps)
```

Now, attach a factor for time of week (weekend/weekday)
```{r}
act_data1$TimeOfWeek <- factor(weekdays(as.Date(act_data1$date)))
levels(act_data1$TimeOfWeek) <- list(Weekend = c("Sunday", "Saturday"), Weekday=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))

head(act_data1)
```

Now calculate and plot avg steps for each 5-min interval for weekdays and weekends.
```{r}
five_min_avg1 <- aggregate(act_data1$steps,  by=list(act_data1$TimeOfWeek, act_data1$interval), FUN=mean)
colnames(five_min_avg1) <- c("ToW", "Interval", "Steps")
```
```{r}
library(lattice)
xyplot(Steps~Interval|ToW, data=five_min_avg1, type="l", layout=c(1,2), main="Comparison of Avg. Steps for Weekends and Weekdays", grid=TRUE)
```

